---
title: "Database sampling"
date: "2017-01-08"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to extrapolatr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Sometimes one wishes to sample (perhaps for quality) a large number of
items to determine that the occurrence probability of unusual cases is
sufficiently low.  Statistically, this is the problem of determining
an unknown binomial probability.  Some efficiencies can be gained by
using a simple group-sequential design, but that does require
computation to determine its operating characteristics.

Herein we suppose that we're sampling a $250000$-item database, and we
want to be sure that the occurrence probability is less than $0.1\%$.
We suppose we'll look at the database after $10000$, $30000$, and
$100000$ items.  After $10000$ items we'll stop sampling if we've seen
$4$ or fewer cases.  After $30000$ items we'll stop if we've seen $19$
or fewer cases.  And after $100000$ items we'll stop if we've see $79$
or fewer cases.

```{r}
steps <- c(10000, 30000, 100000)
crits <- c(4, 19, 79)
```

We can look at the operating characteristics thereof relatively
simply.  We first have a look at the stopping probabilities when the
true event probability is $0.1\%$.

```{r}
library("dbSamplr")

tot.1 <- gsProbs(0.1/100, steps, crits)
tot.1
tot.1 %*% steps + 250000 * (1 - sum(tot.1))
sum(tot.1)
```

Now we'll see what happens if the true event probability is $0.075\%$.

```{r}
tot.2 <- gsProbs(0.075/100, steps, crits)
tot.2
tot.2 %*% steps + 250000 * (1 - sum(tot.2))
sum(tot.2)
```

Now we'll see what happens if the true event probability is $0.05\%$.

```{r}
tot.3 <- gsProbs(0.05/100, steps, crits)
tot.3
tot.3 %*% steps + 250000 * (1 - sum(tot.3))
sum(tot.3)
```

And finally we'll see what happens if the true event probability is
$0.01\%$.

```{r}
tot.4 <- gsProbs(0.01/100, steps, crits)
tot.4
tot.4 %*% steps + 250000 * (1 - sum(tot.4))
sum(tot.4)
```

Some plots are possible.

```{r}
all.res <- cbind(tot.1, tot.2, tot.3, tot.4)

matplot(steps, all.res,
        ylim=c(0, 1), type='l',
	log='x',
	axes=FALSE,
	xlab='Sample point',
	ylab='Stopping probability')
axis(1, at=steps, label=format(steps, scientific=FALSE))
axis(2, las=1)
```